---
title: ICR OPINION WG2
author: Johannes B. Gruber
date: today
format:
  revealjs:
    logo: https://www.opinion-network.eu/img/opinion_i.gif
    footer: "OPINION | Measuring Opinionated Communication"
    smaller: true
    transition: "slide"
    scrollable: true
    embed-resources: true
---

# Introduction

After the first round of annotations (after the pilot), let's see how we are doing in term of intercoder reliability!

1. Look at intercoder agreement
2. Check the preliminary results

```{r setup}
#| include: false
if (!rlang::is_installed("annotinder")) remotes::install_github("ccs-amsterdam/annotinder-r")
library(annotinder)
library(tidyverse); theme_set(hrbrthemes::theme_ipsum_rc())
library(tidycomm)
library(gt)

# if you are me and have the password, load new data, otherwise load it from the repo
if (Sys.getenv("ANNOTINDER_PW") != "") {
  backend_connect(host = "https://cost-tools.up.railway.app",
                  username = "johannesb.gruber@gmail.com",
                  .password = Sys.getenv("ANNOTINDER_PW"))
  data_coded <- download_annotations(20L)
  saveRDS(data_coded, "../data/abstracts_icr-1.rds")
} else {
  data_coded <- readRDS("../data/abstracts_icr-1.rds")
}

abstracts <- readRDS("../data/abstracts.rds")

test_agreement <- function(tbl, title) {
  tbl |> 
    test_icr(unit_var = unit_id, coder_var = coder_id, value,
             cohens_kappa = TRUE,
             fleiss_kappa = TRUE,
             brennan_prediger = TRUE,
             lotus = TRUE,
             s_lotus = TRUE) |> 
    select(where(is.numeric)) |> 
    t() |> 
    tibble::enframe() |> 
    filter(!is.na(value)) |> 
    # mutate(value = str_replace(value, "(\\d\\.\\d{2}).*", "\\1")) |> 
    gt() |> 
    fmt_number(value, decimals = 2L, drop_trailing_zeros = TRUE) |> 
    data_color(
      columns = value,
      method = "numeric",
      palette = c("red", "green"),
      domain = c(0, 1),
      na_color = "#FFF"
    ) |> 
    tab_header(
      title = md(title)
    )
}

disagreement_score <- function(x) {
  code_counts <- table(x)
  proportions <- code_counts / sum(code_counts)
  max(proportions) - sum(proportions[-which.max(proportions)])
}

majority_vote <- function(x) {
  names(which.max(table(x)))
}
```

## Top Coders!

```{r}
coders <- data_coded |> 
  group_by(coder_id) |> 
  mutate(name = value[variable == "welcome"]) |> 
  filter(unit_id != "id") |> 
  summarise(coded = length(unique(unit_id)),
            name = unique(name))

star_coders <- coders |>
  filter(coded == 50L)

coders |> 
  mutate(name = forcats::fct_reorder(name, coded)) |> 
  ggplot(aes(x = coded, y = name, fill = coded)) +
  geom_col() +
  scale_fill_gradient(low = "#A2A4A6", high = "#AE9653") +
  labs(x = NULL, y = NULL, title = "Top Coders!")
```

:::{.fragment .fade-up .absolute top="25%" right="25%"}
<iframe src="https://giphy.com/embed/KcPvrQkRRiM2D75roA" width="459" height="480" frameBorder="0" class="giphy-embed" allowFullScreen></iframe><p><a href="https://giphy.com/gifs/happy-animated-super-star-KcPvrQkRRiM2D75roA"></a></p>
:::

# Measurement Problems {.scrollable}

:::: {.columns}

::: {.column width="50%"}
::: {.incremental}
- Branching in codebook makes calculating intercoder-reliability tough
- What does it mean when somebody answers an abstract *uses a tool* and someone clicked the research does NOT measure opinion (i.e., isn't asked the question)?
- options:
  1. treat *IRRELEVANT* as valid code?
  2. treat as missing and ignore (works only for Krippendorff's $\alpha$)?
  3. calculate agreement on full branch path?
:::
:::
::: {.column width="50%"}
```{r}
disagreement <- data_coded |> 
  filter(variable == "relevant-2",
         coder_id %in% star_coders$coder_id) |> 
  group_by(unit_id) |> 
  mutate(d_score = disagreement_score(value)) |> 
  select(unit_id, coder_id, d_score, value) |> 
  ungroup()

disagreement |> 
  filter(unit_id %in% c("4775", "4567", "1948")) |> 
  select(-d_score) |> 
  pivot_wider(id_cols = "unit_id", names_from = "coder_id", values_from = "value") |> 
  gt()
```
:::

::::

# ICR
## Is this research that measures a human opinion?

:::: {.columns}

::: {.column width="50%"}
```{r}
data_coded |> 
  filter(variable == "relevant-1") |> 
  test_agreement(title = "Agreement (all coders)")
```
:::

::: {.column width="50%"}
```{r}
data_coded |> 
  filter(variable == "relevant-1",
         coder_id %in% star_coders$coder_id) |> 
  test_agreement(title = "Agreement (full cases)")
```
:::

::::

## What understanding or sub-category of opinion is measured here?

:::: {.columns}

::: {.column width="50%"}
```{r}
data_coded |> 
  filter(variable == "concept") |> 
  test_agreement(title = "Agreement (all coders)")
```
:::

::: {.column width="50%"}
```{r}
data_coded |> 
  filter(variable == "concept",
         coder_id %in% star_coders$coder_id) |> 
  test_agreement(title = "Agreement (full cases)")
```
:::

::::

## New sub-categories of opinion?

```{r}

data_coded |> 
  filter(variable == "concept-other",
         value != "IRRELEVANT") |> 
  separate_longer_delim(cols = value, delim = ";") |> 
  count(value, sort = TRUE) |> 
  gt() |> 
  tab_options(container.height = px(500))
```

## Is this research that uses a tool?

:::: {.columns}

::: {.column width="50%"}
```{r}
data_coded |> 
  filter(variable == "relevant-2") |> 
  test_agreement(title = "Agreement (all coders)")
```
:::

::: {.column width="50%"}
```{r}
data_coded |> 
  filter(variable == "relevant-2",
         coder_id %in% star_coders$coder_id) |> 
  test_agreement(title = "Agreement (full cases)")
```
:::

::::

## Disagreement Examples

```{r}
disagreement |> 
  filter(unit_id %in% c("1948", "1017", "597")) |> 
  select(-d_score) |> 
  pivot_wider(id_cols = "unit_id", names_from = "coder_id", values_from = "value") |> 
  gt()
```


```{r}
#| output: asis
#| class: fragment
print_abstract <- function(id) {
  out <- abstracts |> 
    #filter(ID %in% c("1948", "1017", "597"))
    filter(ID == id) |> 
    select(`Article Title`, Abstract)
  cat("ID:", id, "\n\n")
  cat(paste0("**", out$`Article Title`, "**"), "\n\n")
  cat(out$Abstract, "\n\n")
}
print_abstract("1948")
```

```{r}
#| output: asis
#| class: fragment
print_abstract("1017")
```

```{r}
#| output: asis
#| class: fragment
print_abstract("597")
```


## How would you categorize the measurment approach of the tool?

:::: {.columns}

::: {.column width="50%"}
```{r}
data_coded |> 
  filter(variable == "approach") |> 
  test_agreement(title = "Agreement (all coders)")
```
:::

::: {.column width="50%"}
```{r}
data_coded |> 
  filter(variable == "approach",
         coder_id %in% star_coders$coder_id) |> 
  test_agreement(title = "Agreement (full cases)")
```
:::

::::

## New sub-categories of approaches?

```{r}

data_coded |> 
  filter(variable == "approach-other",
         value != "IRRELEVANT") |> 
  separate_longer_delim(cols = value, delim = ";") |> 
  count(value, sort = TRUE) |> 
  gt() |> 
  tab_options(container.height = px(500))
```

## What software application or model or dictionary is employed by the research to measure the concept?

:::: {.columns}

::: {.column width="50%"}
```{r}
data_coded |> 
  filter(variable == "tool") |> 
  test_agreement(title = "Agreement (all coders)")
```
:::

::: {.column width="50%"}
```{r}
data_coded |> 
  filter(variable == "tool",
         coder_id %in% star_coders$coder_id) |> 
  test_agreement(title = "Agreement (full cases)")
```
:::

::::


## Does the abstract mention or hint which data(set) was analysed?

:::: {.columns}

::: {.column width="50%"}
```{r}
data_coded |> 
  filter(variable == "relevant-data") |> 
  test_agreement(title = "Agreement (all coders)")
```
:::

::: {.column width="50%"}
```{r}
data_coded |> 
  filter(variable == "relevant-data",
         coder_id %in% star_coders$coder_id) |> 
  test_agreement(title = "Agreement (full cases)")
```
:::

::::

## What kind of dataset is referenced?

:::: {.columns}

::: {.column width="50%"}
```{r}
data_coded |> 
  filter(variable == "data-kind") |> 
  test_agreement(title = "Agreement (all coders)")
```
:::

::: {.column width="50%"}
```{r}
data_coded |> 
  filter(variable == "data-kind",
         coder_id %in% star_coders$coder_id) |> 
  test_agreement(title = "Agreement (full cases)")
```
:::

::::


## Does the abstract mention or hint at the natural language analysed?

:::: {.columns}

::: {.column width="50%"}
```{r}
data_coded |> 
  filter(variable == "islang") |> 
  test_agreement(title = "Agreement (all coders)")
```
:::

::: {.column width="50%"}
```{r}
data_coded |> 
  filter(variable == "islang",
         coder_id %in% star_coders$coder_id) |> 
  test_agreement(title = "Agreement (full cases)")
```
:::

::::

## Does the abstract mention or hint at the country analysed?

:::: {.columns}

::: {.column width="50%"}
```{r}
data_coded |> 
  filter(variable == "iscountry") |> 
  test_agreement(title = "Agreement (all coders)")
```
:::

::: {.column width="50%"}
```{r}
data_coded |> 
  filter(variable == "iscountry",
         coder_id %in% star_coders$coder_id) |> 
  test_agreement(title = "Agreement (full cases)")
```
:::

::::

# Results
## Valid

```{r}
data_coded |> 
  filter(status == "DONE",
         variable == "relevant-1") |> 
  group_by(unit_id) |> 
  summarise(value = majority_vote(value)) |> 
  count(value) |> 
  ggplot(aes(x = n, y = value)) +
  geom_col(fill = "#042C23") + 
  labs(x = NULL, y = NULL, title = "Is this research that measures a human opinion (however it is defined)?")
```

## Tentative  {.scrollable}

Results are based on a simple majority vote.

```{r}
data_coded |> 
  filter(status == "DONE",
         variable == "concept",
         value != "IRRELEVANT") |> 
  group_by(unit_id) |> 
  summarise(value = majority_vote(value)) |> 
  count(value) |> 
  ggplot(aes(x = n, y = value)) +
  geom_col(fill = "#042C23") + 
  labs(x = NULL, y = NULL, title = "Understanding or sub-category of opinion")
```

```{r}
data_coded |> 
  filter(status == "DONE",
         variable == "relevant-2",
         value != "IRRELEVANT") |> 
  group_by(unit_id) |> 
  summarise(value = majority_vote(value)) |> 
  count(value) |> 
  ggplot(aes(x = n, y = value)) +
  geom_col(fill = "#042C23") + 
  labs(x = NULL, y = NULL, title = "Is this research that uses a tool?")
```

```{r}
data_coded |> 
  filter(status == "DONE",
         variable == "tool",
         value != "IRRELEVANT") |> 
  group_by(unit_id) |> 
  summarise(value = majority_vote(value)) |> 
  count(value) |> 
  ggplot(aes(x = n, y = value)) +
  geom_col(fill = "#042C23") + 
  labs(x = NULL, y = NULL, title = "What software application or model or dictionary?")
```

```{r}
data_coded |> 
  filter(status == "DONE",
         variable == "relevant-data",
         value != "IRRELEVANT") |> 
  group_by(unit_id) |> 
  summarise(value = majority_vote(value)) |> 
  count(value) |> 
  ggplot(aes(x = n, y = value)) +
  geom_col(fill = "#042C23") + 
  labs(x = NULL, y = NULL, title = "mention or hint which data(set)?")
```


```{r}
data_coded |> 
  filter(status == "DONE",
         variable == "data-kind",
         value != "IRRELEVANT") |> 
  group_by(unit_id) |> 
  summarise(value = majority_vote(value)) |> 
  count(value) |> 
  ggplot(aes(x = n, y = value)) +
  geom_col(fill = "#042C23") + 
  labs(x = NULL, y = NULL, title = "What kind of dataset is referenced?")
```


```{r}
data_coded |> 
  filter(status == "DONE",
         variable == "data-source",
         value != "IRRELEVANT") |> 
  group_by(unit_id) |> 
  summarise(value = majority_vote(value)) |> 
  count(value) |> 
  ggplot(aes(x = n, y = value)) +
  geom_col(fill = "#042C23") + 
  labs(x = NULL, y = NULL, title = "Where does the data come from?")
```


```{r}
data_coded |> 
  filter(status == "DONE",
         variable == "islang",
         value != "IRRELEVANT") |> 
  group_by(unit_id) |> 
  summarise(value = majority_vote(value)) |> 
  count(value) |> 
  ggplot(aes(x = n, y = value)) +
  geom_col(fill = "#042C23") + 
  labs(x = NULL, y = NULL, title = "Mentions natural language?")
```

```{r}
data_coded |> 
  filter(status == "DONE",
         variable == "data-lang",
         value != "IRRELEVANT") |> 
  group_by(unit_id) |> 
  summarise(value = majority_vote(value)) |> 
  count(value) |> 
  ggplot(aes(x = n, y = value)) +
  geom_col(fill = "#042C23") + 
  labs(x = NULL, y = NULL, title = "Which natural language?")
```

```{r}
data_coded |> 
  filter(status == "DONE",
         variable == "iscountry",
         value != "IRRELEVANT") |> 
  group_by(unit_id) |> 
  summarise(value = majority_vote(value)) |> 
  count(value) |> 
  ggplot(aes(x = n, y = value)) +
  geom_col(fill = "#042C23") + 
  labs(x = NULL, y = NULL, title = "country analysed?")
```

```{r}
data_coded |> 
  filter(status == "DONE",
         variable == "data-country",
         value != "IRRELEVANT") |> 
  group_by(unit_id) |> 
  summarise(value = majority_vote(value)) |> 
  count(value) |> 
  ggplot(aes(x = n, y = value)) +
  geom_col(fill = "#042C23") + 
  labs(x = NULL, y = NULL, title = "Which natural language?")
```

(Note: the wording of the question was wrong)
