---
title: tool annotation validation
author: Johannes B. Gruber
date: today
format:
  html:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    code-fold: true
    code-tools: true
---

# Introduction

We validate the annotation of tools between our two human coders, and the two human coders plus the LLM.

## Packages and Functions

```{r}
#| include: false
library(googledrive)
library(tidyverse)
library(tidycomm)
library(gt)
library(yardstick)
start <- Sys.time() # note start time for later
setwd(here::here("analysis_visualization"))
```

Function to colour-code ICR results

```{r}
# 𝛼-levels:
#
# 𝛼 > 0.8 reliable
# 𝛼 > 0.667 and 𝛼 < 0.8 tentative conclusions
# 𝛼 < 0.667 not reliable
# (Krippendorff, 2004, pp. 241-243)
icr_print <- function(
  tbl,
  cols = c("Agreement", "Krippendorffs_Alpha")
) {
  alpha_color <- function(x) {
    seg_high <- scales::col_numeric(c("#90EE90", "#006400"), domain = c(0.8, 1)) # light → dark green
    seg_mid <- scales::col_numeric(
      c("#FFA500", "#ADFF2F"),
      domain = c(0.667, 0.8)
    )
    seg_low <- scales::col_numeric(
      c("#8B0000", "#FF6B6B"),
      domain = c(-1, 0.667)
    )

    out <- character(length(x))
    high <- !is.na(x) & x >= 0.8
    mid <- !is.na(x) & x >= 0.667 & x < 0.8
    low <- !is.na(x) & x < 0.667
    out[high] <- seg_high(x[high])
    out[mid] <- seg_mid(x[mid])
    out[low] <- seg_low(x[low])
    out[is.na(x)] <- "#FFFFFF"
    out
  }

  tbl |>
    gt() |>
    fmt_number(
      columns = all_of(cols),
      decimals = 2L,
      drop_trailing_zeros = TRUE
    ) |>
    data_color(
      columns = all_of(cols),
      fn = alpha_color
    )
}

ml_metrics <- metric_set(accuracy, precision, recall, f_meas)
```

## Data

```{r load data}
# get annotation data from G drive
annotation <- "1.2_tools_annotated_manual.xlsx"
if (!file.exists(annotation)) {
  drive_download(
    file = "https://docs.google.com/spreadsheets/d/1a5m1SgQ3ccZN6N1c-u0ZqJkkOKbmYXGjjOcxIefqmf0",
    path = annotation
  )
}

manual_annotation <- map(
  c("validation_DT", "validation_CV", "validation_JG"),
  function(s) {
    rio::import(
      annotation,
      sheet = s
    ) |>
      mutate(annotator = str_remove(s, "validation_"))
  }
) |>
  bind_rows()

llm_annotation <- readRDS("1.2_tools_annotated_long.rds") |>
  select(-request, -cache_file, -query, -annotation) |>
  mutate(annotator = "LLM")
```

Combine data:

```{r}
annotation_df <- bind_rows(
  manual_annotation,
  llm_annotation
) |>
  as_tibble()
```

Only use complete cases:

```{r}
annotation_df <- annotation_df |>
  complete(annotator, tool) |>
  group_by(tool) |>
  mutate(
    missing = rowSums(across(
      c(
        category,
        task_representation,
        task_classification,
        task_target,
        software
      ),
      is.na
    ))
  ) |>
  filter(all(missing == 0)) |>
  ungroup()
```

# Main

Below are the annotation columns we measure reliability for:

```{r}
# Columns to compare
compare_cols <- c(
  "category",
  "task_representation",
  "task_classification",
  "task_target",
  "software"
)
```

## Human ICR

```{r}
#| tbl-cap: "Intercoder reliability for human annotators"
annotation_df |>
  filter(annotator != "LLM") |>
  test_icr(
    unit_var = tool,
    coder_var = annotator,
    all_of(compare_cols),
    agreement = TRUE,
    holsti = FALSE,
    kripp_alpha = TRUE,
    cohens_kappa = FALSE,
    fleiss_kappa = FALSE,
    brennan_prediger = FALSE,
    lotus = FALSE,
    s_lotus = FALSE,
    na.omit = TRUE
  ) |>
  icr_print()
```

## human + LLM ICR

```{r}
#| tbl-cap: "Intercoder reliability for human+LLM annotators"
annotation_df |>
  test_icr(
    unit_var = tool,
    coder_var = annotator,
    all_of(compare_cols),
    cohens_kappa = FALSE,
    fleiss_kappa = FALSE,
    brennan_prediger = FALSE,
    lotus = FALSE,
    s_lotus = FALSE,
    holsti = FALSE,
    na.omit = TRUE
  ) |>
  icr_print(cols = c("Agreement", "Krippendorffs_Alpha"))
```

# Machine learning metrics

Alternatively to the ICR metrics, we can also calculate more classic machine learning metrics: precision, recall and F1 scores.
Since we have two annotators, I calculate these measures for both of them separately:

```{r}
annotation_df_wide <- annotation_df |>
  pivot_longer(
    cols = all_of(compare_cols),
    names_to = "variable",
    values_to = "annotation"
  ) |>
  pivot_wider(
    id_cols = tool,
    names_from = c(annotator, variable),
    values_from = annotation
  ) |>
  mutate(
    across(contains("_category"), function(x) {
      factor(
        x,
        levels = c(
          "algorithm",
          "approach",
          "commercial_api",
          "data_collection",
          "general_library",
          "general_purpose_nlp",
          "gui_tool",
          "infrastructure",
          "linguistic_resource",
          "model",
          "programming_language",
          "single_purpose_nlp"
        )
      )
    }),
    across(c(contains("_task"), contains("_software")), function(x) {
      factor(
        x,
        levels = c(
          "yes",
          "no"
        )
      )
    })
  )
```

### Daniel vs the machine:

```{r}
coder <- "DT"
map(compare_cols, function(col) {
  annotation_df_wide |>
    ml_metrics(
      truth = paste0(coder, "_", col),
      estimate = paste0("LLM_", col)
    ) |>
    add_column(variable = col)
}) |>
  bind_rows() |>
  pivot_wider(
    id_cols = variable,
    names_from = .metric,
    values_from = .estimate
  ) |>
  icr_print(cols = c("accuracy", "precision", "recall", "f_meas"))
```


### Christina vs the machine

```{r}
coder <- "CV"
map(compare_cols, function(col) {
  annotation_df_wide |>
    ml_metrics(
      truth = paste0(coder, "_", col),
      estimate = paste0("LLM_", col)
    ) |>
    add_column(variable = col)
}) |>
  bind_rows() |>
  pivot_wider(
    id_cols = variable,
    names_from = .metric,
    values_from = .estimate
  ) |>
  icr_print(cols = c("accuracy", "precision", "recall", "f_meas"))
```

### Johannes vs the machine

```{r}
coder <- "JG"
map(compare_cols, function(col) {
  annotation_df_wide |>
    ml_metrics(
      truth = paste0(coder, "_", col),
      estimate = paste0("LLM_", col)
    ) |>
    add_column(variable = col)
}) |>
  bind_rows() |>
  pivot_wider(
    id_cols = variable,
    names_from = .metric,
    values_from = .estimate
  ) |>
  icr_print(cols = c("accuracy", "precision", "recall", "f_meas"))
```

### Majority vs the machine

```{r}
# row-wise majority vote across the three human coders.
# with 3 coders a binary (yes/no) variable always has a majority; for the
# multi-class category a tie (all three disagree) returns NA and is dropped
# from the metrics below.
majority_vote <- function(...) {
  vals <- as.character(c(...))
  vals <- vals[!is.na(vals)]
  if (length(vals) == 0) {
    return(NA_character_)
  }
  tab <- sort(table(vals), decreasing = TRUE)
  if (length(tab) > 1 && tab[[1]] == tab[[2]]) {
    return(NA_character_)
  }
  names(tab)[1]
}

annotation_df_wide_maj <- annotation_df_wide |>
  rowwise() |>
  mutate(
    MAJ_category = majority_vote(DT_category, CV_category, JG_category),
    MAJ_task_representation = majority_vote(
      DT_task_representation,
      CV_task_representation,
      JG_task_representation
    ),
    MAJ_task_classification = majority_vote(
      DT_task_classification,
      CV_task_classification,
      JG_task_classification
    ),
    MAJ_task_target = majority_vote(
      DT_task_target,
      CV_task_target,
      JG_task_target
    ),
    MAJ_software = majority_vote(DT_software, CV_software, JG_software)
  ) |>
  ungroup() |>
  mutate(
    MAJ_category = factor(MAJ_category, levels = levels(DT_category)),
    across(
      c(
        MAJ_task_representation,
        MAJ_task_classification,
        MAJ_task_target,
        MAJ_software
      ),
      function(x) factor(x, levels = c("yes", "no"))
    )
  )
```

```{r}
coder <- "MAJ"
majority_metrics <- map(compare_cols, function(col) {
  annotation_df_wide_maj |>
    ml_metrics(
      truth = paste0(coder, "_", col),
      estimate = paste0("LLM_", col)
    ) |>
    add_column(variable = col)
}) |>
  bind_rows() |>
  pivot_wider(id_cols = variable, names_from = .metric, values_from = .estimate)

majority_metrics |>
  icr_print(cols = c("accuracy", "precision", "recall", "f_meas"))
```


## Confusion Matrices

Software:

```{r}
conf_mat(
  annotation_df_wide_maj,
  truth = "MAJ_software",
  estimate = "LLM_software"
)
```

task_representation:

```{r}
conf_mat(
  annotation_df_wide_maj,
  truth = "MAJ_task_representation",
  estimate = "LLM_task_representation"
)
```

task_classification:

```{r}
conf_mat(
  annotation_df_wide_maj,
  truth = "MAJ_task_classification",
  estimate = "LLM_task_classification"
)
```

task_target:

```{r}
conf_mat(
  annotation_df_wide_maj,
  truth = "MAJ_task_target",
  estimate = "LLM_task_target"
)
```


# Disagreement

```{r}
disagreement <- map(compare_cols, function(col) {
  annotation_df |>
    group_by(tool) |>
    filter(n_distinct(.data[[col]]) > 1) |>
    ungroup() |>
    select(tool, all_of(col), annotator) |>
    pivot_wider(
      id_cols = tool,
      names_from = annotator,
      values_from = all_of(col)
    )
}) |>
  set_names(compare_cols)

rio::export(disagreement, "disagreement.xlsx")
```


# wrap up

Save important result:

```{r}
saveRDS(majority_metrics, "1.3_humans-LLM-validation.rds")
write_csv(majority_metrics, "1.3_humans-LLM-validation.csv")
```

Save data if it takes long to preprocess it here. 
Afterwards we get some information which is important to reproduce the report.

```{r}
sessionInfo()
Sys.time()
# note how long the script takes to (re-)run
Sys.time() - start
```
