---
title: "PDF annotation"
format:
  pdf:
    keep-tex: true
knitr:
  opts_chunk:
    dev: pdf
    echo: false
    fig.path: "fig/"
---

# Intro

Tables and plots to answer our research questions:

- RQ1a: Which theoretical concepts are being measured in opinion mining?
- RQ1b: Which methods and measurements are used for opinion mining?
- RQ1c: Which tools are used for opinion mining?
- RQ1d: Which data is used in CSS opinion mining?
- RQ2: Are there disciplinary differences in concepts, methods, tools, and data in opinion mining?
- RQ3: Do concepts, methods, tools, and data in opinion mining change over time?

## Packages and functions

```{r setup}
#| output: false
setwd(here::here("analysis_visualization"))
library(tidyverse)
theme_set(theme_minimal())
start <- Sys.time() # note start time for later
```

## Data

```{r}
results <- rio::import("2._annotation-results.csv")
abstracts_clean <- rio::import(
  "https://github.com/JBGruber/opinion-wg2/raw/main/relevance-prediction/Opinion%20Papers%20Abstracts%20-%20data_abstracts_clean_withpredictedrelevance.csv"
)
tool_dict <- rio::import("1.2_tool_dict.csv")
tools_df <- results |>
  filter(
    variable %in% c("Q1_1_Tool-Name", "Q2_1_Tool-Name"),
    !result %in% c("-", "IRRELEVANT"),
    nchar(result) > 1
  ) |>
  separate_longer_delim(cols = result, delim = "; ") |>
  rename(tool = result) |>
  left_join(tool_dict, by = "tool")
tools_annotated <- readRDS("1.2_tools_annotated_long.rds")
```

# Main
## Validation


## RQ1a theoretical concepts

Question Q4_1_Opinion-Evaluation

```{r}
# Define categories for opinion evaluation concepts
categories_evaluations_l <- jsonlite::read_json("2._concept_dict.json")

categories_evaluations_dict <- tibble(
  category = names(categories_evaluations_l),
  concept = categories_evaluations_l
) |>
  unnest_longer(concept)

# Process and categorize concepts
concepts <- results |>
  filter(
    variable == "Q4_1_Opinion-Evaluation",
    !result %in% c("-", "IRRELEVANT")
  ) |>
  separate_longer_delim(cols = result, delim = "; ") |>
  mutate(
    result = trimws(result),
    result_lower = tolower(result)
  ) |>
  left_join(categories_evaluations_dict, by = c("result_lower" = "concept"))

# Aggregate counts by category
concepts_aggregated <- concepts |>
  count(category, name = "n") |>
  arrange(desc(n))
```

```{r}
#| fig-width: 7
#| fig-height: 5
#| fig-cap: "Categories of Opinion Evaluation Concepts"
concepts_aggregated |>
  filter(category != "coding-error") |>
  mutate(category = fct_reorder(category, n)) |>
  ggplot(aes(x = n, y = category)) +
  geom_col() +
  labs(
    title = "What theoretical concepts are being measured?",
    x = "Count",
    y = NULL
  )
```

## RQ1b methods and measurements used for opinion mining?

```{r}
#| fig-width: 15
#| fig-height: 7
#| fig-cap: "Approaches used for Opinion Measurement"
approaches <- results |>
  filter(variable == "Q3_1_Approach", !result %in% c("-", "IRRELEVANT")) |>
  separate_longer_delim(cols = result, delim = ";") |>
  mutate(
    result = trimws(result),
    result = case_when(
      str_detect(
        result,
        "Deep|Transformers|Embedding"
      ) ~ "Deep learning/Embeddings, Transformers etc",
      str_detect(
        result,
        "[U|u]nsupervised"
      ) ~ "Classic Unsupervised Machine Learning",
      str_detect(
        result,
        "\\b[S|s]upervised|[M|a]chine|Ensemble|Class"
      ) ~ "Classic Supervised Machine Learning",
      str_detect(result, "[D|d]ict|[L|l]exi") ~ "Dictionaries",
      str_detect(result, "Proprietary") ~ "Proprietary and unknown to us",

      TRUE ~ "Other"
    )
  )

approaches |>
  count(result) |>
  mutate(approach = fct_reorder(result, n)) |>
  ggplot(aes(x = n, y = approach)) +
  geom_col() +
  labs(x = NULL, y = NULL)
```

## RQ1d: Which data is used in CSS opinion mining?


## Target  



```{r}
#| fig-width: 7
#| fig-height: 7
results |>
  filter(
    variable == "Q3_2_Target-specific-Measurement",
    !result %in% c("-", "IRRELEVANT")
  ) |>
  count(result) |>
  mutate(result = fct_reorder(result, n)) |>
  ggplot(aes(x = n, y = result)) +
  geom_col() +
  labs(
    title = "Does the computational approach consider specific targets?",
    x = NULL,
    y = NULL
  )
```

## Validation

```{r}
#| fig-width: 7
#| fig-height: 7
results |>
  filter(variable == "Q3_3_Validation", !result %in% c("-", "IRRELEVANT")) |>
  count(result) |>
  mutate(result = fct_reorder(result, n)) |>
  ggplot(aes(x = n, y = result)) +
  geom_col() +
  labs(
    title = "Do the authors report validation of their measurement?",
    x = NULL,
    y = NULL
  )
```

## RQ2: disciplinary differences

### concepts

### methods

### tools

### data


## RQ3: change over time

### concepts


### methods

```{r}
#| fig-width: 9
#| fig-height: 4
approaches |>
  filter(
    result %in%
      c(
        "Classic Supervised Machine Learning",
        "Deep learning/Embeddings, Transformers etc",
        "Dictionaries"
      )
  ) |>
  left_join(abstracts_clean, by = "id") |>
  count(result, year) |>
  group_by(year) |>
  mutate(pct = n / sum(n)) |>
  ggplot(aes(x = year, y = pct, colour = result)) +
  geom_line() +
  scale_color_viridis_d() +
  scale_y_continuous(labels = scales::percent) +
  theme(legend.position = "bottom") +
  labs(
    title = "Approaches used for Opinion Measurement",
    x = NULL,
    y = NULL,
    colour = NULL
  )
```

### tools


### data

# wrap up

Save data if it takes long to preprocess it here. 
Afterwards we get some information which is important to reproduce the report.

```{r}
sessionInfo()
Sys.time()
# note how long the script takes to (re-)run
Sys.time() - start
```

